---
title: "VE303-Ph2-AnalysisWorkflow to run LME and stats models on VE303 phase2 data"
author: "Rajita M"
date: "2024-08-26"
output: html_document
---

## Workflow to assess VE303-002 data for 79 patient IA
## Note packages and source files here
```{r, warning=FALSE, include=FALSE}

library(plyr)
library(openxlsx)
library(RColorBrewer)
library(tidyverse)
library(vegan)
library(here)
library(ggrepel)
library(ggpubr)
library(scales)
library(Maaslin2)
library(ZIBR)
library(nlme)
library(lme4)
library(dplyr)

source("colors.R") #source colors

```

## Print R version and working directory.
```{r}
#sessionInfo()
here()

```

## Set up directory structure for output - this will create a new directory with the current date to save results to
```{r, include=FALSE}
results <- paste(Sys.Date(), "results dose-response", sep = "_")
dir.create(here(results), showWarnings = TRUE)

# Function Will be needed for survival analysis ##
show.me <- function(x){
  if(as.data.frame(x[, c("rec.diagnosis")])[,1][1] == "Bad"){x$rec.day.in.treatment.surv <- x$rec.day.in.treatment.surv}
    if(as.data.frame(x[, c("rec.diagnosis")])[,1][1] == "Good"){x$rec.day.in.treatment.surv <- x$Day.in.treatment  }

  return(x)}

```

## Load required files
```{r, include=FALSE}


## Import cleaned metadata Based on merged SPR
dnameta = read_csv(here("Input files/2022-03-08 VE303_dna_metadata_80IA_mult.csv"))

#### Import updated  marker panel with targeted relative abundance estimation : TO BE COMBINED with metadata file
ve303 = read_csv(here("Input files/2022-01-11 VE303_Ph2_Extended_Marker_Data.csv"))

## Import cleaned recurrence metadata
rec = read_csv(here("Input files/2022-03-09 CDI_Recurrences_IA2_treatment_assignment-OUT.csv"))

## Import cleaned conmeds metadata
conmed = read_csv(here("Input files/2021-10-05 VE303-002_ConMeds_treatment_assignment-OUT.csv"))


## Import primary vs secondary CDI population metadata
study_pops = read_csv(here("Input files/2022-03-16 VE303 prior CDI summary.csv")) %>%
  rename(Subject.Number = Subject)

```

## Merge Extended marker panel with metadata to create VE303 subject level data file
## Merge VE303 subject level data and ConMed data to establish time of abx dosing and interaction with recurrence
## Merge VE303 subject level data and recurrence data 

```{r}
analysis_78 <- "yes"

## Uses files imported above ##
source("VE303.data.import.merge.R")


conduct_abx <- "Abx"
ve303.dnameta[is.na(ve303.dnameta$Abx.group_joined_M),]["Abx.group_joined_M"] <- "Missing"
if(conduct_abx == "Abx"){ve303.dnameta <- ve303.dnameta %>% filter(!Abx.group_joined_M %in% c("Missing"  ) )}


```


```{r}

ve303.dnameta_preabx <- ve303.dnameta %>% filter(collect.pre.post.abx == "Pre.abx")  
ve303.dnameta_allabx <- ve303.dnameta 

taxplot <- "Pre Abx"

if(taxplot == "Pre Abx"){
  #Add frequency variable for all rows - this gives every detection/non-det result a value of 1
  ve303.dnameta_plot <- data.frame(ve303.dnameta_preabx, freq = 1)
}

if(taxplot == "All Abx"){
  #Add frequency variable for all rows - this gives every detection/non-det result a value of 1
  ve303.dnameta_plot <- data.frame(ve303.dnameta_allabx, freq = 1)
}
 


##To plot zeroes
ve303.dnameta_plot.alt = ve303.dnameta_plot %>%
  mutate(est_relative_abundance_panel_alt = if_else(detection_status != "Detected", 0, est_relative_abundance_panel)) %>%
  group_by(Subject.Number, Visit.Name ) %>%
  mutate(.,est_relative_abundance_panel_alt_sum = sum(est_relative_abundance_panel_alt)) %>%
  ungroup()



```

## Plot VE303 detection and calculate summary stats  - All DETECTED events Pre-Abx OR Pre- and Post-Abx  depending on selection in script
#Add frequency variable for all rows - this gives every detection/non-det result a value of 1

```{r, include=FALSE}

source("VE303.tables.summary.stats.R")

```

### VE303 Abundance - only detected strains
```{r}


##To plot zeroes
ve303.dnameta.det.alt = ve303.dnameta_plot %>%
  mutate(est_relative_abundance_panel_alt = if_else(detection_status != "Detected", 0, est_relative_abundance_panel)) %>%
  group_by(Subject.Number, Visit.Name ) %>%
  mutate(.,est_relative_abundance_panel_alt_sum = sum(est_relative_abundance_panel_alt)) %>%
  ungroup()


# Filter the data frame to only include VE303 Strain Detected by the marker panel
ve303.dnameta.det <- ve303.dnameta_plot %>% 
  filter(detection_status == "Detected")

# Filter the data frame to include VE303 Strain Detected AND Insufficient data by the marker panel

ve303.dnameta.det_ID <- ve303.dnameta_plot %>% 
  filter(detection_status %in% c("Detected", "Insufficient data"))


ve303.dnameta.det.tp <- ve303.dnameta.det %>% 
  arrange(Subject.Number, Visit.Name.Norm) %>% 
  group_by(Visit.Name.Norm, organism) %>%      # Select a single representative sample from each Subject for each Visit.Name.Norm & result in the same df as above
  distinct(Subject.Number, .keep_all = TRUE)

 

```

# VE303 Abundance summary by Subject and timepoint, dfs for plotting
```{r,include=FALSE}

source("VE303.abundance.summary.wrangling.R")




```

 

```{r}

analysis_type <- 'marker_panel'


```

 

```{r}

##############################################################################
# Prepare full VE303 array for analysis
##############################################################################

d1 <- ve303.dnameta.det.alt

d1 <- d1 %>%
  rename(
    rel_abund = est_relative_abundance_panel_alt
    )

VE303_sp <- (unique(d1$organism))
 
d11.total <- d1
days_numeric <- as.numeric(d11.total$Day.in.treatment.norm)
days_numeric_adjusted <- as.numeric(plyr::mapvalues(d11.total$Day.in.treatment.norm, from= c("Screening","-11"  ,"5" ,"10", "12" , "13" ), to=c(-11,-11 ,7,14 ,14,14)))

batch_factor <- plyr::mapvalues(d11.total$Seq.batch, from= c("Batch 1", "Batch 2", "Batch 3", "Batch 4"), to=c(1, 2, 3, 4))
response_factor <- plyr::mapvalues(d11.total$rec.diagnosis, from= c("Bad", "Good"), to=c(1, 2))
response_factor_Wk8 <- plyr::mapvalues(d11.total$rec.diagnosis.Wk8, from= c("Bad", "Good"), to=c(1, 2))
response_factor_joined <- plyr::mapvalues(d11.total$collect.pre.post.rec_joined, from= c("Post.rec", "Pre.rec"), to=c(1, 2))
cohort_factor <- plyr::mapvalues(d11.total$TRT.norm, from= c("Placebo","VE303 Low Dose","VE303 High Dose"), to=c(1, 2,3))
Abx_factor <- as.factor(d11.total$Abx.group_joined_M)

d11.total$Timepoint <- as.numeric(days_numeric)
d11.total$Timepoint_adj <- as.numeric(days_numeric_adjusted)
d11.total$Response <- as.numeric(response_factor)
d11.total$Response_Wk8 <- as.numeric(response_factor_Wk8)
d11.total$Response_joined <- as.numeric(response_factor_joined)
d11.total$Treatment <- as.numeric(cohort_factor)
d11.total$Batch <- as.numeric(batch_factor)
d11.total$Abx <-  (Abx_factor)



VE303_sp.all <- d11.total
 

# ##############################################################################
# # Prepare sum VE303 array for ZIBR analysis
# ##############################################################################
 
d1.total <- ve303.abund.subj.total.time.det.alt

d1.total <- d1.total %>%
 rename(
    ve303_tot = est_relative_abundance_panel_alt_sum,
    ve303_tot_prop = prop_det_alt_sum
     )

d11.total <- d1.total

days_numeric <- as.numeric(d11.total$Day.in.treatment)
days_numeric_adjusted <- as.numeric(plyr::mapvalues(d11.total$Day.in.treatment, from= c("Screening","-11"  ,"5" ,"10", "12" , "13" ), to=c(-11,-11 ,7,14 ,14,14)))


batch_factor <- plyr::mapvalues(d11.total$Seq.batch, from= c("Batch 1", "Batch 2", "Batch 3", "Batch 4"), to=c(1, 2,3,4))
response_factor <- plyr::mapvalues(d11.total$rec.diagnosis, from= c("Bad", "Good"), to=c(1, 2))
response_factor_Wk8 <- plyr::mapvalues(d11.total$rec.diagnosis.Wk8, from= c("Bad", "Good"), to=c(1, 2))
response_factor_joined <- plyr::mapvalues(d11.total$collect.pre.post.rec_joined, from= c("Post.rec", "Pre.rec"), to=c(1, 2))
cohort_factor <- plyr::mapvalues(d11.total$TRT.norm, from= c("Placebo","VE303 Low Dose","VE303 High Dose"), to=c(1, 2,3))
Abx_factor <- as.factor(d11.total$Abx.group_joined_M)


d11.total$Timepoint <- (days_numeric)
d11.total$Timepoint_adj <- (days_numeric_adjusted)
d11.total$Response <- as.numeric(response_factor)
d11.total$Response_Wk8 <- as.numeric(response_factor_Wk8)
d11.total$Response_joined <- as.numeric(response_factor_joined)
d11.total$Treatment <- as.numeric(cohort_factor)
d11.total$Batch <- as.numeric(batch_factor)
d11.total$Abx <-  (Abx_factor)


VE303_sp.total <- d11.total
 

colpick.total <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Response_joined","Batch" ,"Visit.Name", "Treatment", "ve303_tot", "ve303_tot_prop")
 
colpick.all <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Response_joined","Batch" ,"Visit.Name", "Treatment","rel_abund", "organism")

if(conduct_abx == "Abx"){
  colpick.total <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Response_joined","Batch" ,"Visit.Name", "Treatment", "Abx", "ve303_tot", "ve303_tot_prop")
 
colpick.all <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Response_joined","Batch" ,"Visit.Name", "Treatment", "Abx" ,"rel_abund", "organism")
   }


```
 

# Function to find subjects with required timepoints ##
```{r}

find.subjects <- function(subs, dtest, tt_choose){
  subsub <- c() 
  for (sub1 in subs) {
  dtt <- dtest %>% filter(Subject.Number == sub1)
  times <- dtt$Timepoint_adj %>% unique()
  if(setequal(intersect(tt_choose,times), tt_choose)){
    subsub <- c(subsub, sub1)} }
  return(subsub)}

```


#### Including desired timepoints ####

### ttest for single timepoint ###
```{r}

### Identify subjects to include, remove subs where not all timepoints available

#tt_choose <- c(1,7,14, 28)
daylabel <- "D14"
tt_choose <- c(14)

dtest <- d11.total
subjects <- dtest$Subject.Number %>% unique()
subsub <- find.subjects(subs = subjects, dtest = d11.total, tt_choose = tt_choose)
print(paste0("Subjects included : " ,subsub %>% length() ))

#############################

d11.total <- VE303_sp.total

VE303_sp1.total <- d11.total[(d11.total$Timepoint_adj %in% tt_choose),][colpick.total]
VE303_sp1.total <- VE303_sp1.total[ (VE303_sp1.total$Sample.ID != "BK1494623"),][colpick.total]
VE303_sp1.total <- VE303_sp1.total[(VE303_sp1.total$Subject.Number) %in% subsub,]

#VE303_sp1.total %>% filter(Treatment  == 2 & Response_Wk8 == 1) %>% 
#  distinct(Subject.Number, .keep_all = TRUE)

 
d11.total <- VE303_sp.all
VE303_sp1.all <- d11.total[(d11.total$Timepoint_adj %in% tt_choose),][colpick.all]
VE303_sp1.all <- VE303_sp1.all[ (VE303_sp1.all$Sample.ID != "BK1494623"),][colpick.all]
VE303_sp1.all <- VE303_sp1.all[(VE303_sp1.all$Subject.Number) %in% subsub,]

#VE303_sp1.all %>% filter(Treatment  == 2 & Response_Wk8 == 1) %>% 
#  distinct(Subject.Number, .keep_all = TRUE)
data_by_subject <- split(VE303_sp1.total, VE303_sp1.total$Subject.Number)




```
# Nonparametric unpaired Wilcox test and t-test comparing VE303 abundance and VE303 proportion across treatment groups at Day 14
```{r, include=FALSE}

####################################################################################################

## As this analysis is for a single timepoint, and there's a duplicate from an unscheduled event in the LD group
VE303_sp1.total  <- VE303_sp1.total %>% distinct(Subject.Number, .keep_all = TRUE)

VE303_sp1.total_LD <- VE303_sp1.total %>% filter(Treatment == "2") 
VE303_sp1.total_HD <- VE303_sp1.total %>% filter(Treatment == "3")  
VE303_sp1.total_PL <- VE303_sp1.total %>% filter(Treatment == "1")  

#FUNCTION
compare_cohort_ttest <- function(V1, V2, comparison){
  
all_p <- c()
all_stat <- c()

tt_VE303tot_abun <- 
  t.test(log10(V1$ve303_tot + 0.00001), log10(V2$ve303_tot + 0.00001))
pv_abun_tot <- tt_VE303tot_abun$p.value
stat_abun_tot <- tt_VE303tot_abun$statistic


tt_VE303tot_prop <- t.test( (V1$ve303_tot_prop ),  (V2$ve303_tot_prop ))
pv_prop_tot <- tt_VE303tot_prop$p.value
stat_prop_tot <- tt_VE303tot_prop$statistic

tt_VE303tot_prev  <- t.test(sign(V1$ve303_tot),sign(V2$ve303_tot))
pv_prev_tot <- tt_VE303tot_prev$p.value
stat_prev_tot <- tt_VE303tot_prev$statistic
ttest_df <- data.frame(comparison,pv_abun_tot, stat_abun_tot, pv_prop_tot, 
                       stat_prop_tot, pv_prev_tot, stat_prev_tot)
names(ttest_df) <- c("Comparison","p-value Abundance", "t-Statistic Abundance",
                    "p-value Proportion", "t-Statistic Proportion",
                    "p-value Prevalence", "t-Statistic Prevalence")
return(ttest_df)
}

compare_cohort_wilcox <- function(V1, V2, comparison){
all_p <- c()
all_stat <- c()
tt_VE303tot_abun <- wilcox.test(log10(V1$ve303_tot + 0.00001), 
                                log10(V2$ve303_tot + 0.00001) , conf.int = TRUE)
pv_abun_tot <- tt_VE303tot_abun$p.value
stat_abun_tot <- tt_VE303tot_abun$estimate


tt_VE303tot_prop <- wilcox.test( (V1$ve303_tot_prop ),  (V2$ve303_tot_prop ), conf.int = TRUE)
pv_prop_tot <- tt_VE303tot_prop$p.value
stat_prop_tot <- tt_VE303tot_prop$estimate

tt_VE303tot_prev  <- wilcox.test(sign(V1$ve303_tot),sign(V2$ve303_tot), conf.int = TRUE)
pv_prev_tot <- tt_VE303tot_prev$p.value
stat_prev_tot <- tt_VE303tot_prev$estimate

wilcox_df <- data.frame(comparison,pv_abun_tot, stat_abun_tot, pv_prop_tot, 
                       stat_prop_tot, pv_prev_tot, stat_prev_tot)

names(wilcox_df) <- c("Comparison","p-value Abundance", "t-Statistic Abundance",
                    "p-value Proportion", "t-Statistic Proportion",
                    "p-value Prevalence", "t-Statistic Prevalence")
return(wilcox_df)
}

result_LD_HD <- compare_cohort_ttest(VE303_sp1.total_LD, VE303_sp1.total_HD, "LD vs HD")
result_PL_HD <- compare_cohort_ttest(VE303_sp1.total_PL, VE303_sp1.total_HD, "Placebo vs HD")
result_PL_LD <- compare_cohort_ttest(VE303_sp1.total_PL, VE303_sp1.total_LD, "Placebo vs LD")
result_all_t <- rbind(result_LD_HD, result_PL_HD, result_PL_LD)

result_all_t$`p-adjust Abundance` <-  p.adjust( (result_all_t$`p-value Abundance`), method = "BH")
result_all_t$`p-adjust Proportion` <-  p.adjust( (result_all_t$`p-value Proportion`), method = "BH")
result_all_t$`p-adjust Prevalence` <-  p.adjust( (result_all_t$`p-value Prevalence`), method = "BH")


result_LD_HD <- compare_cohort_wilcox(VE303_sp1.total_LD, VE303_sp1.total_HD, "LD vs HD")
result_PL_HD <- compare_cohort_wilcox(VE303_sp1.total_PL, VE303_sp1.total_HD, "Placebo vs HD")
result_PL_LD <- compare_cohort_wilcox(VE303_sp1.total_PL, VE303_sp1.total_LD, "Placebo vs LD")

result_all_wil <- rbind(result_LD_HD, result_PL_HD, result_PL_LD)

result_all_wil$`p-adjust Abundance` <-  p.adjust( (result_all_wil$`p-value Abundance`), method = "BH")
result_all_wil$`p-adjust Proportion` <-  p.adjust( (result_all_wil$`p-value Proportion`), method = "BH")
result_all_wil$`p-adjust Prevalence` <-  p.adjust( (result_all_wil$`p-value Prevalence`), method = "BH")


 
write.xlsx(result_all_wil,  file = here(results, paste(Sys.Date(), taxplot, daylabel,"Total VE303 Wilcox_Abun_Prop_Table S1.xlsx", sep = " ")) )


```


 


# Functions to loop over VE303 individual species at Day 14 for Wilcoxon unpaired test and ttest
```{r include=FALSE}

### FUNCTION
compare_cohorts_species_ttest <- function(V1, V1_tot, V2, V2_tot, specs){
  tt_VE303tot_abun <- t.test(log10(V1_tot$ve303_tot + 0.00001), log10(V2_tot$ve303_tot + 0.00001)) 
pv_abun_tot <- tt_VE303tot_abun$p.value
stat_abun_tot <- tt_VE303tot_abun$statistic

tt_VE303tot_prop <- t.test( (V1_tot$ve303_tot_prop ),  (V2_tot$ve303_tot_prop ))
pv_prop_tot <- tt_VE303tot_prop$p.value
stat_prop_tot <- tt_VE303tot_prop$statistic

pv_prev_tot <- NA
stat_prev_tot <- NA

  for (spec1 in specs) {
    t_prev <- t.test(V1[V1$organism == spec1, ]$Ndet , V2[V2$organism == spec1, ]$Ndet )
pv_prev <- c(pv_prev, t_prev$p.value)
stat_prev <- c(stat_prev, t_prev$statistic)

t_abun <- t.test(log10(0.00001 + V1[V1$organism == spec1, ]$rel_abund) , log10(0.00001 + V2[V2$organism == spec1, ]$rel_abund))
pv_abun <- c(pv_abun, t_abun$p.value)
stat_abun <- c(stat_abun, t_abun$statistic)  }

pv_abun <- c(pv_abun, pv_abun_tot, pv_prop_tot)
stat_abun <- c(stat_abun, stat_abun_tot, stat_prop_tot)
qv_abun <- p.adjust(pv_abun, method = "BH")

pv_prev <- c(pv_prev, pv_prev_tot)
qv_prev <- p.adjust(pv_prev, method = "BH")
v_dum <- NA

stat_prev <- c(stat_prev, stat_prev_tot, v_dum)
pv_prev <- c(pv_prev,v_dum)
qv_prev <- c(qv_prev, v_dum)

specs_all <- c(specs, "Total VE303", "Total VE303 Proportion")
ttest_df <- data.frame(specs_all, pv_abun,qv_abun,stat_abun , pv_prev,qv_prev, stat_prev)
names(ttest_df) <- c("Species", "p-value Abundance","Adjusted p-value Abundance", "t-Statistic Abundance", "p-value Prevalence", "Adjusted p-value Prevalence", "t-Statistic Prevalence")
return(ttest_df) }

### FUNCTION
compare_cohorts_species_wilcoxtest <- function(V1, V1_tot, V2, V2_tot, specs){
  tt_VE303tot_abun <- wilcox.test(log10(V1_tot$ve303_tot + 0.00001), log10(V2_tot$ve303_tot + 0.00001),conf.int = TRUE) 
  pv_abun_tot <- tt_VE303tot_abun$p.value
stat_abun_tot <- tt_VE303tot_abun$estimate

tt_VE303tot_prop <- wilcox.test( (V1_tot$ve303_tot_prop ),  (V2_tot$ve303_tot_prop ), conf.int = TRUE)
pv_prop_tot <- tt_VE303tot_prop$p.value
stat_prop_tot <- tt_VE303tot_prop$estimate 

pv_prev_tot <- NA
stat_prev_tot <- NA

for (spec1 in specs) {
    t_prev <- wilcox.test(V1[V1$organism == spec1, ]$Ndet , V2[V2$organism == spec1, ]$Ndet , conf.int = TRUE)
pv_prev <- c(pv_prev, t_prev$p.value)
stat_prev <- c(stat_prev, t_prev$estimate)

t_abun <- wilcox.test(log10(0.00001 + V1[V1$organism == spec1, ]$rel_abund) , log10(0.00001 + V2[V2$organism == spec1, ]$rel_abund), conf.int = TRUE)
pv_abun <- c(pv_abun, t_abun$p.value)
stat_abun <- c(stat_abun, t_abun$estimate)  }

pv_abun <- c(pv_abun, pv_abun_tot, pv_prop_tot)
stat_abun <- c(stat_abun, stat_abun_tot, stat_prop_tot)
qv_abun <- p.adjust(pv_abun, method = "BH")

pv_prev <- c(pv_prev, pv_prev_tot)
qv_prev <- p.adjust(pv_prev, method = "BH")
v_dum <- NA

stat_prev <- c(stat_prev, stat_prev_tot, v_dum)
pv_prev <- c(pv_prev,v_dum)
qv_prev <- c(qv_prev, v_dum)

specs_all <- c(specs, "Total VE303", "Total VE303 Proportion")
wilcox_df <- data.frame(specs_all, pv_abun,qv_abun,stat_abun , pv_prev,qv_prev, stat_prev)
names(wilcox_df) <- c("Species", "p-value Abundance","Adjusted p-value Abundance", "t-Statistic Abundance", "p-value Prevalence", "Adjusted p-value Prevalence", "t-Statistic Prevalence")
return(wilcox_df) }


```

 

### Loop over day 14 log transformed ttest for all strains : HD vs LD ###
```{r}

VE303_sp1.all_PL <- VE303_sp1.all %>% filter(Treatment == "1") 
VE303_sp1.all_LD <- VE303_sp1.all %>% filter(Treatment == "2") 
VE303_sp1.all_HD <- VE303_sp1.all %>% filter(Treatment == "3")

VE303_sp1.all_PL$Ndet  <- sign(VE303_sp1.all_PL$rel_abund)
VE303_sp1.all_LD$Ndet  <- sign(VE303_sp1.all_LD$rel_abund)
VE303_sp1.all_HD$Ndet  <- sign(VE303_sp1.all_HD$rel_abund)


pv_abun <- c()
stat_abun <- c()

pv_prev <- c()
stat_prev <- c()

V1 <- VE303_sp1.all_LD
V1_tot <- VE303_sp1.total_LD
V2 <- VE303_sp1.all_HD
V2_tot <-VE303_sp1.total_HD
#VE303_sp1.total_PL

specs <- unique(VE303_sp.all$organism)
test_type <- "ttest"

ttest_all <- compare_cohorts_species_ttest(V1, V1_tot, V2, V2_tot, specs)
wiltest_all <- compare_cohorts_species_wilcoxtest(V1, V1_tot, V2, V2_tot, specs)

write.xlsx(ttest_all,  file = here(results, paste(Sys.Date(), taxplot, daylabel,"Species TTest_Abun_Prev_Table S1.xlsx", sep = " ")) )

ttest_all


```
  

##Choose which response metadata was given. Assign recurrence_type: Response_Wk8, Response_joined (pre/post), Response OR if model was run without Response metadata i.e recurrence_type = "Response Excluded"
 
### FITTING LME / MAASLIN TO INCLUDE SUBJECTS WITH MISSING SAMPLES

```{r}

##############################################################################
# Prepare full VE303 array for MAASLIN analysis
##############################################################################

d1 <- ve303.dnameta.det.alt %>% filter(collect.pre.post.abx == "Pre.abx")

if(conduct_abx == "Abx"){
 if(taxplot == "Pre Abx"){d1 <- ve303.dnameta.det.alt %>% filter(collect.pre.post.abx == "Pre.abx")}
 if(taxplot == "All Abx"){d1 <- ve303.dnameta.det.alt  }
}
 

d1 <- d1 %>%
  rename(
    rel_abund = est_relative_abundance_panel
    )

d1.total <- ve303.abund.subj.total.time.det.alt %>% filter(collect.pre.post.abx == "Pre.abx")

if(conduct_abx == "Abx"){
 if(taxplot == "Pre Abx"){d1.total <- ve303.abund.subj.total.time.det.alt %>% filter(collect.pre.post.abx == "Pre.abx")}
 if(taxplot == "All Abx"){d1.total <- ve303.abund.subj.total.time.det.alt  }
}


d1.total <- d1.total %>%
 rename(
    ve303_tot = est_relative_abundance_panel_alt_sum,
    ve303_tot_prop = prop_det_alt_sum
     )


VE303_sp <- (unique(d1$organism))
 
d11.total <- d1

days_numeric <- as.numeric(d11.total$Day.in.treatment.norm)
days_numeric_adjusted <- as.numeric(plyr::mapvalues(d11.total$Day.in.treatment.norm, from= c("Screening","-11"  ,"5" ,"10", "12" , "13" ), to=c(-11,-11 ,7,14 ,14,14)))

  
#  if (including_NAs == "yes"){
#    d11.total$Wk.8.Response[is.na(d11.total$Wk.8.Response)] <- "PD"}

batch_factor <- plyr::mapvalues(d11.total$Seq.batch, from= c("Batch 1", "Batch 2", "Batch 3", "Batch 4"), to=c(1, 2 ,3, 4))
response_factor <- plyr::mapvalues(d11.total$rec.diagnosis, from= c("Bad", "Good"), to=c(1, 2))
response_factor_Wk8 <- plyr::mapvalues(d11.total$rec.diagnosis.Wk8, from= c("Bad", "Good"), to=c(1, 2))
cohort_factor <- plyr::mapvalues(d11.total$TRT.norm, from= c("Placebo","VE303 Low Dose","VE303 High Dose"), to=c(1, 2,3))
response_factor_joined <- plyr::mapvalues(d11.total$collect.pre.post.rec_joined, from= c("Post.rec", "Pre.rec"), to=c(1, 2))
Abx_factor <-  d11.total$Abx.group_joined_M


d11.total$Timepoint <- as.numeric(days_numeric)
d11.total$Timepoint_adj <- as.numeric(days_numeric_adjusted)
d11.total$Response <- as.numeric(response_factor)
d11.total$Response_Wk8 <- as.numeric(response_factor_Wk8)
d11.total$Response_joined <- as.numeric(response_factor_joined)
d11.total$Treatment <- as.numeric(cohort_factor)
d11.total$Batch <- as.numeric(batch_factor)
d11.total$Abx <- as.factor(Abx_factor)

d11.total$rec.TRT =  paste((d11.total$TRT.norm ),d11.total$rec.diagnosis.Wk8)

d11.total$rec.TRT.combined =  if_else(d11.total$rec.TRT %in% c("VE303 High Dose Good", "VE303 Low Dose Good"), "Dosed Good",
                                      if_else(d11.total$rec.TRT %in% c("VE303 High Dose Bad", "VE303 Low Dose Bad"), "Dosed Bad",d11.total$rec.TRT))

 

VE303_sp.all <- d11.total



# ##############################################################################
# # Prepare sum VE303 array for Maaslin analysis
# ##############################################################################
 
d1.total <- ve303.abund.subj.total.time.det.alt  %>% filter(collect.pre.post.abx == "Pre.abx")

d1.total <- d1.total %>%
 rename(
    ve303_tot = est_relative_abundance_panel_alt_sum
     )

d11.total <- d1.total

days_numeric <- as.numeric(d11.total$Day.in.treatment)
days_numeric_adjusted <- as.numeric(plyr::mapvalues(d11.total$Day.in.treatment, from= c("Screening","-11"  ,"5" ,"10", "12" , "13" ), to=c(-11,-11 ,7,14 ,14,14)))


batch_factor <- plyr::mapvalues(d11.total$Seq.batch, from= c("Batch 1", "Batch 2", "Batch 3"), to=c(1, 2, 3))
response_factor <- plyr::mapvalues(d11.total$rec.diagnosis, from= c("Bad", "Good"), to=c(1, 2))
response_factor_Wk8 <- plyr::mapvalues(d11.total$rec.diagnosis.Wk8, from= c("Bad", "Good"), to=c(1, 2))
cohort_factor <- plyr::mapvalues(d11.total$TRT.norm, from= c("Placebo","VE303 Low Dose","VE303 High Dose"), to=c(1, 2,3))
response_factor_joined <- plyr::mapvalues(d11.total$collect.pre.post.rec_joined, from= c("Post.rec", "Pre.rec"), to=c(1, 2))
Abx_factor <-  d11.total$Abx.group_joined_M


d11.total$Timepoint <- (days_numeric)
d11.total$Timepoint_adj <- (days_numeric_adjusted)
d11.total$Response <- as.numeric(response_factor)
d11.total$Response_Wk8 <- as.numeric(response_factor_Wk8)
d11.total$Response_joined <- as.numeric(response_factor_joined)
d11.total$Treatment <- as.numeric(cohort_factor)
d11.total$Batch <- as.numeric(batch_factor)
d11.total$Abx <- as.factor(Abx_factor)

d11.total$rec.TRT =  paste(d11.total$TRT.norm,d11.total$rec.diagnosis.Wk8)
d11.total$rec.TRT.combined =  if_else(d11.total$rec.TRT %in% c("VE303 High Dose Good", "VE303 Low Dose Good"), "Dosed Good",
                                      if_else(d11.total$rec.TRT %in% c("VE303 High Dose Bad", "VE303 Low Dose Bad"), "Dosed Bad",d11.total$rec.TRT))


VE303_sp.total <- d11.total


colpick.total <- c("Sample.ID","Subject.Number","Timepoint_adj","Response", "Response_Wk8","Batch" ,"Visit.Name", "Treatment", "ve303_tot", "ve303_tot_prop","collect.pre.post.rec")
 
colpick.all <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Batch" ,"Visit.Name", "Treatment","rel_abund", "organism", "collect.pre.post.rec")

if(conduct_abx == "Abx"){
  colpick.total <- c("Sample.ID","Subject.Number","Timepoint_adj","Response", "Response_Wk8","Batch" ,"Visit.Name", "Treatment", "Abx.group_joined_M","ve303_tot", "ve303_tot_prop","collect.pre.post.rec")
 
colpick.all <- c("Sample.ID","Subject.Number","Timepoint_adj","Response","Response_Wk8","Batch" ,"Visit.Name", "Treatment",  "Abx.group_joined_M","rel_abund", "organism", "collect.pre.post.rec")
}


```


```{r}
 

# VE303 Abundance summary by Subject and timepoint, df for plotting INCLUDING zeros
VE303_sp.all_Maaslin <- VE303_sp.all %>% 
  select(Sample.ID,Subject.Number, Timepoint_adj ,rec.diagnosis, rec.diagnosis.Wk8, collect.pre.post.rec,Batch, rec.TRT, rec.TRT.combined, Visit.Name,  organism, TRT.norm, rel_abund ) %>%
  spread(organism, rel_abund, fill=0) %>% # This fills in zeros for all events not detected - needed to calc mean
  left_join(., distinct(ve303.dnameta[c("Sample.ID","shannon_diversity")]), by = "Sample.ID") 

colpick.all <- c("Subject.Number","Timepoint_adj","rec.diagnosis", "rec.diagnosis.Wk8", "rec.TRT", "rec.TRT.combined","Batch" ,"Visit.Name", "TRT.norm", "collect.pre.post.rec")

if(conduct_abx == "Abx"){
  VE303_sp.all_Maaslin <- VE303_sp.all %>% 
  select(Sample.ID,Subject.Number, Timepoint_adj ,rec.diagnosis, rec.diagnosis.Wk8, collect.pre.post.rec,Batch, Abx.group_joined_M, rec.TRT, rec.TRT.combined, Visit.Name,  organism, TRT.norm, rel_abund ) %>%
  spread(organism, rel_abund, fill=0) %>% # This fills in zeros for all events not detected - needed to calc mean
  left_join(., distinct(ve303.dnameta[c("Sample.ID","shannon_diversity")]), by = "Sample.ID") 
 
colpick.all <- c("Subject.Number","Timepoint_adj","rec.diagnosis", "rec.diagnosis.Wk8", "rec.TRT", "rec.TRT.combined","Batch",  "Abx.group_joined_M" ,"Visit.Name", "TRT.norm", "collect.pre.post.rec")  
}


VE303_sp.all_Maaslin

```

```{r}

prepMaaslin <- function(VE303_sp.all_Maaslin){
  
  ### MODEL 0 ####
  # ARRAY FOR ALL TREATMENTS,  KEEPING ALL TIMEPOINTS
  VE303_sp.all_Maaslin_All <- VE303_sp.all_Maaslin 
  
  ### MODEL 1 ####
  # ARRAY FOR ALL TREATMENTS,  KEEPING ALL TIMEPOINTS PRE-RECURRENCE 
  VE303_sp.all_Maaslin_postrecfiltered <- VE303_sp.all_Maaslin %>% filter(collect.pre.post.rec != "Post.rec")

  ### MODEL 2 ####
  # ARRAY FOR ALL TREATMENTS, ONLY KEEPING TIMEPOINTS THAT ARE BOTH PRE-RECURRENCE +++ EARLY TIMEPOINTS UP TO DAY 28 ONLY
  timefilt <- c("Day 1","Day 7","Day 14","Day 28")
  VE303_sp.all_Maaslin_time_postrecfiltered_D28 <- VE303_sp.all_Maaslin %>% filter((collect.pre.post.rec != "Post.rec") & (Visit.Name %in% timefilt))
  
  ### MODEL 3 ####
  # ARRAY FOR ALL TREATMENTS, ONLY KEEPING TIMEPOINTS THAT ARE BOTH PRE-RECURRENCE +++ EARLY TIMEPOINTS UP TO DAY 14 ONLY
  timefilt1 <- c( "Day 1","Day 7","Day 14")
  VE303_sp.all_Maaslin_time_postrecfiltered_D14 <- VE303_sp.all_Maaslin %>% filter((collect.pre.post.rec != "Post.rec") & (Visit.Name %in% timefilt1))

  ### MODEL 4 ####
  # ARRAY FOR ALL TREATMENTS, ONLY KEEPING EARLY UP TO DAY 28 REGARDLESS OF RECURRENCE STATUS
  VE303_sp.all_Maaslin_time_filtered <- VE303_sp.all_Maaslin %>% filter((Visit.Name %in% timefilt))


  ### MODEL 5 ####
  # ARRAY FOR ALL TREATMENTS, ONLY KEEPING EARLY UP TO DAY 14 REGARDLESS OF RECURRENCE STATUS
  VE303_sp.all_Maaslin_time_filtered_D14 <- VE303_sp.all_Maaslin %>% filter((Visit.Name %in% timefilt1))
  
  
  ### MODEL 6 ####
  # ARRAY FOR ALL TREATMENTS, ONLY KEEPING DAY 14 REGARDLESS OF RECURRENCE STATUS
  timefilt2 <- c("Day 14")
  VE303_sp.all_Maaslin_time_filtered_D14_only <- VE303_sp.all_Maaslin %>% filter((Visit.Name %in% timefilt2))
  
  ### MODEL 7 ####
  # ARRAY FOR ALL TREATMENTS, KEEPING DAY 7 and later timepoints + Pre-recurrence  
  timefilt2 <- c("Day 14")
  VE303_sp.all_Maaslin_time_postrecfiltered_D1Onward <- VE303_sp.all_Maaslin %>% filter((collect.pre.post.rec != "Post.rec") & (Visit.Name != "Day 1"))
  
  models <- c("Model 0","Model 1", "Model 2", "Model 3", "Model 4", "Model 5", "Model 6", "Model 7" )
  model_description <- c("AllTime","Exclude post-recurrence", "Exclude post-recurrence_D28", "Exclude post-recurrence_D14",
                         "Exclude post-D28",  "Exclude post-D14",  "Include D14 Only",  "Exclude Day1 and post-recurrence" )
  Arrays <- list(VE303_sp.all_Maaslin_All,VE303_sp.all_Maaslin_postrecfiltered, VE303_sp.all_Maaslin_time_postrecfiltered_D28, VE303_sp.all_Maaslin_time_postrecfiltered_D14, VE303_sp.all_Maaslin_time_filtered,VE303_sp.all_Maaslin_time_filtered_D14, VE303_sp.all_Maaslin_time_filtered_D14_only, VE303_sp.all_Maaslin_time_postrecfiltered_D1Onward)
  Output <- list(Arrays, models, model_description)
  return(Output)
}

```

# FILTER TIMEPOINTS POST DAY 63 and post abx

```{r}

VE303_sp.all_Maaslin <- VE303_sp.all_Maaslin %>% filter(Timepoint_adj <= 63 & Timepoint_adj > 0)
 

```

# CHOOSE WHETHER TO DO TREATMENT COMPARISON FOR ALL LEVELS OR JUST WITH HD LD. 
```{r}
# Run LME to compare HD vs placebo and LD vs placebo
treats <- "Dosed and Placebo"

# Run LME to compare HD vs LD 
#treats <- "Dosed"

```

```{r, include=FALSE}

if(treats == "Dosed and Placebo"){
Arr_out <- prepMaaslin(VE303_sp.all_Maaslin)
Arrays <- Arr_out[[1]]
models <- Arr_out[[2]]
model_description <- Arr_out[[3]]
}

# ARRAY FOR DOSED COMPARISON HD VS LD ONLY, ALL TIMEPOINTS #
if(treats == "Dosed"){
VE303_sp.all_Maaslin_filtered <- VE303_sp.all_Maaslin %>% filter(TRT.norm != "Placebo")
Arr_out <- prepMaaslin(VE303_sp.all_Maaslin_filtered)
Arrays <- Arr_out[[1]]
models <- Arr_out[[2]]
model_description <- Arr_out[[3]]
}

 

```

 
##CHOOSE Whether Model runs on final recurrence status OR Week 8 Recurrence status
```{r}
recurrence_type <- "rec.diagnosis.Wk8"
## Can take values "treatment","treatment response","interaction treatment response", "response", "prepost response"

#if(conduct_abx != "Abx"){
#  modelling <- "prepost response"
#  }

#modelling <- "prepost response"
modelling <- "treatment Abx"

## When conduct_abx = "Abx", can take value "treatment Abx" 
#if(conduct_abx == "Abx"){modelling <- "treatment Abx"}

param_choice <- "LOG"
if(param_choice == "AST"){params <- c("NONE", "AST")}
if(param_choice == "LOG"){params <- c("NONE","LOG")}
if(param_choice == "CLR"){params <- c("CLR", "NONE")}
 
modelling

```


```{r}

run_maaslin <- "Yes"

print("the models available to run including different data cuts, pre/post recurrence and pre/post antibiotics samples ~")
model_description

print("The code in the next chunk runs the model using data up to Day 14, and excluding samples collected after a recurrence")

```





```{r, include=FALSE}


if(run_maaslin == "Yes"){ 
  countvec <- c(4) # Choose the model that uses data up to Day 14, and excluding samples collected after a recurrence
#  countvec <- c(1,2,3,4,5,6)

if(modelling == "treatment Abx"){
    for (cvec in countvec) {  
    VE303_sp.all_Maaslin_test <- Arrays[[cvec]]
    model_type <- model_description[cvec]
  
    VE303_sp_div <- c(VE303_sp, "shannon_diversity")
    ve303_organism <- VE303_sp.all_Maaslin_test[VE303_sp]
    ve303_organism$VE303_total <- rowSums(ve303_organism)
    pseudo <- 0.00001
    if(params[2] == "LOG"){ve303_organism == pseudo + ve303_organism}

    ve303_organism$shannon_diversity <- VE303_sp.all_Maaslin_test$shannon_diversity
    ve303_metadata <- VE303_sp.all_Maaslin_test[colpick.all]

    ve303_organism <- data.frame(ve303_organism)
    rownames(ve303_organism) <- rownames(ve303_organism)

    ve303_metadata <- data.frame(ve303_metadata)
    rownames(ve303_metadata) <- rownames(ve303_metadata)
    
      if(treats == "Dosed"){fixed <- c("TRT.norm", "Batch", "Abx.group_joined_M")
                         ref_level <- c(  "Batch,1")} 
      if(treats == "Dosed and Placebo"){fixed <- c("TRT.norm", "Batch", "Abx.group_joined_M")
        ref_level <- c(  "TRT.norm,Placebo","Batch,1")} 
  
    
    fit_data2 = Maaslin2(
    input_data = ve303_organism, 
    input_metadata = ve303_metadata, 

    normalization =  params[1],
    transform =  params[2],
    output = paste(Sys.Date(), treats, taxplot, modelling,  param_choice,model_type,sep = " "), 
    
    fixed_effects = fixed,
    reference =  ref_level,
    random_effects = c("Subject.Number"))
      }
    }
  }



```
 
# load model results from Maaslin tsv output and wrangle to view in excel
```{r}

#treats <- "Dosed"
treats <- "Dosed and Placebo"

if(conduct_abx == "Abx"){
  if(taxplot == "Pre Abx"){
  model_type <- "Exclude post-recurrence_D14"
  location <- paste(Sys.Date(), treats, taxplot,modelling, "LOG", model_type,sep = " ")
  location_data <- paste(location ,"all_results.tsv",sep = "/")}

  model_result <- read_tsv(here(location_data))
  models_list <- list()
  
  for (x1 in (model_result$metadata %>% unique())) {
    model_result1 <- model_result %>% 
      filter(metadata == x1)
    
    model_result2 <- model_result1[
      with(model_result1, order(value, pval)),] %>%
      filter(pval < 1.0)
    models_list <-  append(models_list, list(model_result2)) }
  
  models_1 <- lapply(models_list, function(x) cbind(" "=rownames(x), x))
  ## Writes to multi sheet excel file ###
  writexl::write_xlsx(models_1, here(location, paste("results", ".xlsx", sep="")) )
  writexl::write_xlsx(models_1, here(results, paste(treats ,"LME Table S1.xlsx", sep=" ")) )
}
 

```
   
 
 
 

 
