---
title: "VE303-Ph2-AnalysisWorkflow to assign engraftment categories and create Time To Adverse Event Curve."
author: "Rajita M"
date: "2024/08/22"
output: html_document
---

## Workflow to assess VE303-002 data for 79 patients
## Note packages and source files here
```{r, warning=FALSE, include=FALSE}


library(haven)
library(openxlsx)
library(RColorBrewer)
library(tidyverse)
library(vegan)
library(here)
library(ggrepel)
library(ggpubr)
library(scales)
library(nlme)
library(lme4)
library(survminer)
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ggfortify)
library(sandwich)
library(msm)
library(writexl)
library(broom)
library(PupillometryR)

source("colors.R") #source colors


```

## Print R version and working directory.
```{r}
#sessionInfo()
here()
```

## Set up directory structure for output - this will create a new directory with the current date to save results to
```{r, include=FALSE}
results <- paste(Sys.Date(), "TimeToEvent", sep = "_")
dir.create(here(results), showWarnings = TRUE)


show.me <- function(x){
  if(as.data.frame(x[, c("rec.diagnosis")])[,1][1] == "Bad"){x$rec.day.in.treatment.surv <- x$rec.day.in.treatment.surv}
  if(as.data.frame(x[, c("rec.diagnosis")])[,1][1] == "Good"){x$rec.day.in.treatment.surv <- max(x$Day.in.treatment)}
  return(x)}


calc_prop_7or14 <- function(a){
  if( ( (c("Day 7") %in% a$Visit.Name.Norm) & (!c("Day 14") %in% a$Visit.Name.Norm))){a$prop_D7orD14  <- a[a["Visit.Name.Norm"] == "Day 7",]$Day.7} 
  if( (c("Day 14") %in% a$Visit.Name.Norm) ){a$prop_D7orD14  <- a[a["Visit.Name.Norm"] == "Day 14",]$Day.14} 
  if(( (!c("Day 7") %in% a$Visit.Name.Norm) & (!c("Day 14") %in% a$Visit.Name.Norm)) ){a$prop_D7orD14  <- NA} 
                          return(a)}
                


```

## Load required files
```{r, include=FALSE}
## Import cleaned metadata Based on merged SPR

dnameta = read_csv(here("Input files/2022-03-08 VE303_dna_metadata_80IA_mult.csv"))

#### Import updated  marker panel with targeted relative abundance estimation : TO BE COMBINED with metadata file
ve303 = read_csv(here("Input files/2022-01-11 VE303_Ph2_Extended_Marker_Data.csv"))

## Import cleaned recurrence metadata

## Final recurrence list with subject 103106 as non-recurrent
rec = read_csv(here("Input files/2022-03-09 CDI_Recurrences_IA2_treatment_assignment-OUT.csv"))

## Import cleaned conmeds metadata
conmed = read_csv(here("Input files/2021-10-05 VE303-002_ConMeds_treatment_assignment-OUT.csv"))


## Import primary vs secondary CDI population metadata
study_pops = read_csv(here("Input files/2022-03-16 VE303 prior CDI summary.csv")) %>%
  rename(Subject.Number = Subject)

#### Import simpson and inverse simpson index per sample calculated in endogenous taxa processing notebook in PD R markdown file
simpson = read.csv(here("Input files/2024-02-19 Species_Simpson_Indices.csv"))

adtte <- read_xpt("Input files/adtte.xpt")  %>% 
  filter(PARAM =="Time to CDI Recurrence, Sensitivity 2") %>% 
  mutate( Subject.Number = as.numeric(SUBJID)) %>%
  select( Subject.Number,AVAL, CNSR) %>%
  rename(time_rec = AVAL,
         status_rec = CNSR)

adsl <- read_xpt("Input files/adsl.xpt")

adtte_mod <- adtte %>% left_join(dnameta %>% 
                                   select(Subject.Number, TRT.norm) %>%
                                   distinct(Subject.Number, .keep_all = TRUE), by = "Subject.Number")

adtte_mod
```

 

 


## Merge Extended marker panel with metadata to create VE303 subject level data file
## Merge VE303 subject level data and ConMed data to establish time of abx dosing and interaction with recurrence
## Merge VE303 subject level data and recurrence data 

```{r}
analysis_78 <- "yes"

## Uses files imported above ##
source("VE303.data.import.merge.R")

ve303.dnameta <- ve303.dnameta %>% left_join(simpson, by = "MGS.Sample.ID")  

```

 

 

```{r}

ve303.dnameta_preabx <- ve303.dnameta %>% filter(collect.pre.post.abx == "Pre.abx")  
ve303.dnameta_allabx <- ve303.dnameta 

taxplot <- "Pre Abx"

if(taxplot == "Pre Abx"){
  #Add frequency variable for all rows - this gives every detection/non-det result a value of 1
  ve303.dnameta_plot <- data.frame(ve303.dnameta_preabx, freq = 1)
}

if(taxplot == "All Abx"){
  #Add frequency variable for all rows - this gives every detection/non-det result a value of 1
  ve303.dnameta_plot <- data.frame(ve303.dnameta_allabx, freq = 1)
}

ve303.dnameta_allabx %>% distinct(Subject.Number, .keep_all = TRUE) %>% filter(TRT.norm != "Placebo")

ve303.dnameta_allabx %>% filter(Visit.Name == "Day 14") %>% distinct(Subject.Number, .keep_all = TRUE) %>% filter(TRT.norm != "Placebo") 

ve303.dnameta_preabx %>% filter(Visit.Name == "Day 14") %>% distinct(Subject.Number, .keep_all = TRUE) %>% filter(TRT.norm != "Placebo")


##To plot zeroes
ve303.dnameta_plot.alt = ve303.dnameta_plot %>%
  mutate(est_relative_abundance_panel_alt = if_else(detection_status != "Detected", 0, est_relative_abundance_panel)) %>%
  group_by(Subject.Number, Visit.Name ) %>%
  mutate(.,est_relative_abundance_panel_alt_sum = sum(est_relative_abundance_panel_alt)) %>%
  ungroup()

```

 

```{r}



event_type <- "Recurrence"

####################################################################################################
ve303_names <- ve303.dnameta_plot$organism %>% unique()
plotting_tax <- c( "Proportion VE303", "VE303-08")

# Uncomment below 2 lines to generate results for all strains -  cox model results for Table S3
#plotting_tax1 <- c( "Proportion VE303")
#plotting_tax <- c(plotting_tax1, ve303_names)


plotting_coh <- c("Dosed" )
quantiles <- c("Median" )

whichmedian <- function(x) min(which(x > median(x)))

#Choose day up to which engraftment will be calculated
day_engraft <- 16
day_engraft_start <- 12


day_engraft_label <- "D14 only"
prop_label <- "Separate Days"

HD_KMtrt      <- c(29,29,26,25, 25, 25, 25, 25, 24, 24, 24, 24, 24, 24, 24, 24)
LD_KMtrt      <- c(27,27,25,22, 20, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17)
Placebo_KMtrt <- c(22,22,21,20, 19, 17, 16, 16, 16, 14, 13, 13, 13, 12, 12, 12)

KMtrt_labels <- data.frame(HD_KMtrt, LD_KMtrt, Placebo_KMtrt)
names(KMtrt_labels) <- c("VE303 High Dose", "VE303 Low Dose", "Placebo")

Delta_labels = paste0("Delta = ", c(0.1, 0.2, 0.3))
# A data frame with labels for each facet
f_labels <- data.frame(TRT.norm = c("Placebo", "VE303 High Dose", "VE303 Low Dose"), 
                       Delta_Annotate = Delta_labels)


```





#Code Below will generate KM and Cox analysis for 2-category High/Low Engraftment based on Median and Third quantiles 
# AND 4-category Engraftment based on first - fourth quantiles 
```{r}
plot_tax <- "init"
 
for (quantile in quantiles) {
  for (plot_tax in plotting_tax) {
    print(plot_tax)
    for (plot_coh in plotting_coh) {
      source("KM exposure recurrence quartile loops.R")
    }
  }
  
}


```
 

## Plot VE303 detection and calculate summary stats  - All DETECTED events Pre-Abx OR Pre- and Post-Abx  depending on selection in script
#Add frequency variable for all rows - this gives every detection/non-det result a value of 1

```{r, include=FALSE}

source("VE303.tables.summary.stats.R")


```

## Plots VE303 colonization Figure 1A, Extended data figure 2A,B,C, Extended data Figure 9, 
## Plots Age in recurrent vs non-recurrent dosed subjects: Extended data Figure 5B

```{r, include=FALSE}


ve303.dnameta <- ve303.dnameta_plot.alt
source("plot.detection.allmark.R")


```



## Create tabular summary of VE303 strain detection at scheduled timepoints
## Table S2
```{r}

CD_summary <- ve303.dnameta %>% filter(Visit.Name != "UNSCHED") %>% 
  mutate(detection_binary = if_else(detection_status == "Detected", 1, 0) ) %>%
  select(TRT.norm, Subject.Number, organism, Visit.Name, detection_binary) %>%
  pivot_wider(names_from = Visit.Name, values_from = detection_binary) %>%
  arrange(TRT.norm, organism) 
#  arrange(TRT.norm, Subject.Number) 

col_order <- c("TRT.norm", "Subject.Number" ,"organism" , "Screening", "Day 1" , "Day 7" ,
                "Day 14" , "Day 28" , "Day 56", "Day 168"  )

CD_summary1 <- CD_summary[, col_order]

write.csv(CD_summary1, file = here(results, paste(Sys.Date(), taxplot, "Table S2_Collection_Detection_Summary.csv", sep = " ")), row.names = FALSE)

```

### VE303 Abundance - only detected strains
```{r, include=FALSE}

##To plot zeroes
ve303.dnameta.det.alt = ve303.dnameta_plot %>%
  mutate(est_relative_abundance_panel_alt = if_else(detection_status != "Detected", 0, est_relative_abundance_panel)) %>%
  group_by(Subject.Number, Visit.Name ) %>%
  mutate(.,est_relative_abundance_panel_alt_sum = sum(est_relative_abundance_panel_alt)) %>%
  ungroup()

ve303.dnameta.det.alt$rec.TRT1 =  paste((ve303.dnameta.det.alt$TRT.norm ),ve303.dnameta.det.alt$rec.diagnosis.Wk8)

ve303.dnameta.det.alt$rec.TRT <- factor(ve303.dnameta.det.alt$rec.TRT1,levels = c("Placebo Bad" , "Placebo Good","VE303 High Dose Bad" ,"VE303 High Dose Good",  "VE303 Low Dose Bad","VE303 Low Dose Good"))


ve303.dnameta.det.alt$rec.TRT.combined =  if_else(ve303.dnameta.det.alt$rec.TRT1 %in% c("VE303 High Dose Good", "VE303 Low Dose Good"), "VE303 Dosed Good", if_else(ve303.dnameta.det.alt$rec.TRT1 %in% c("VE303 High Dose Bad", "VE303 Low Dose Bad"), "VE303 Dosed Bad",ve303.dnameta.det.alt$rec.TRT1))

 
source("VE303.abundance.summary.wrangling.R")

# data frame used for diversity LME analyses # 
vvtest5 <- ve303.dnameta.det.alt %>%
  distinct(Subject.Number, Visit.Name.Norm, .keep_all = TRUE) 
days_long <- c("Screening", "Day 1"  , "Day 7" , "Day 14" , "Day 28", "Day 56" , "Day 168")

 
```
 


## Plots for exposure to individual strains up to Day 14 across cohorts
# Figure 1D
```{r}

exposures_relative <- ve303.dnameta.det.alt %>% 
  filter(Visit.Name  %in%  c( "Day 7"  ,   "Day 14"  ) ) %>%  
  #filter(Visit.Name  %in%  c( "Day 7"  ,   "Day 14"  ,  "Day 28") ) %>%  
  select(Subject.Number, organism, TRT.norm, Visit.Name.Norm, Visit.Name,rec.diagnosis.Wk8, 
         est_relative_abundance_panel_alt, MGS.Sample.ID) %>%
  group_by(Subject.Number, TRT.norm, Visit.Name.Norm) %>% 
  spread(organism, est_relative_abundance_panel_alt, fill=0) %>% # This fills in zeros for all events not detected - needed to calc mean
  gather(organism, est_relative_abundance_panel_alt, 7: ncol(.)) %>% 
  group_by(Subject.Number, organism) %>% 
  mutate(subject_exposure_perstrain = mean(est_relative_abundance_panel_alt)) %>%
  ungroup() %>%
  distinct(Subject.Number, organism, .keep_all = TRUE) %>%  
  select(Subject.Number, organism, TRT.norm,  rec.diagnosis.Wk8, subject_exposure_perstrain ) %>%
  mutate(subject_exposure_binary = if_else(subject_exposure_perstrain>0, 1, 0))
  
  
##############

exposure_prevalence <-  exposures_relative %>%
  group_by(organism, TRT.norm) %>%
  mutate(fraction_exposed = sum(subject_exposure_binary)/n_distinct(Subject.Number)) %>%
  ungroup()  %>%
  select(organism, TRT.norm, fraction_exposed) %>%
  distinct(organism, TRT.norm, .keep_all = TRUE)


exposures <- exposure_prevalence
exposures$TRT.norm = factor(exposures$TRT.norm,
                                         levels = c("Placebo", "VE303 Low Dose"  , "VE303 High Dose"))
textsize <- 16

panel1D <- ggplot(exposures %>% filter(TRT.norm != "Placebo"), 
       aes(x=organism, y=((fraction_exposed*100) ), fill=TRT.norm, color=TRT.norm)) +

  geom_bar(stat="identity", alpha = 0.7, width = 0.5,position=position_dodge(width = 0.5) )+
  theme_bw() +
  #facet_grid(.~organism, scales = "free_x", space = "free") +
  scale_fill_manual(values = coh.cols) +
  scale_color_manual(values = coh.cols) +
  theme(axis.text.x = element_text(size = textsize, angle = 45, hjust = 1, vjust =1), 
        axis.title.x = element_text(size = textsize ),
        plot.title = element_text(size = textsize ),
        axis.text.y = element_text(size = textsize, hjust = 1, vjust =0.5) ,     
        axis.title.y = element_text(size = textsize, hjust = 0.5, vjust = 1) ,
        strip.text = element_text(size = textsize-3), legend.title = element_text(size = textsize),
        legend.text = element_text(size = textsize -1)) +
  #scale_x_discrete(labels = c("Placebo" = "Placebo" , "VE303 Low Dose" = "VE303\nLow Dose",
  #                            "VE303 High Dose" = "VE303\nHigh Dose" )) +

    labs(title= "", y="Strain prevalence, %", x="", 
       color="Treatment", fill="Treatment" )
ggsave(filename = here(results, paste(Sys.Date(), taxplot,"VE303 strain prevalence_Figure1D.pdf", sep = " ")), 
       width = 13, height = 5, dpi = 800 )
##############
panel1D

```
 
## Plots for exposure to individual strains up to Day 28 across cohorts
# Extended data Figure 2D
```{r}

exposures_relative <- ve303.dnameta.det.alt %>% 
  #filter(Visit.Name  %in%  c( "Day 7"  ,   "Day 14"  ) ) %>%  
  filter(Visit.Name  %in%  c( "Day 7"  ,   "Day 14"  ,  "Day 28") ) %>%  
  select(Subject.Number, organism, TRT.norm, Visit.Name.Norm, Visit.Name,rec.diagnosis.Wk8, 
         est_relative_abundance_panel_alt, MGS.Sample.ID) %>%
  group_by(Subject.Number, TRT.norm, Visit.Name.Norm) %>% 
  spread(organism, est_relative_abundance_panel_alt, fill=0) %>% # This fills in zeros for all events not detected - needed to calc mean
  gather(organism, est_relative_abundance_panel_alt, 7: ncol(.)) %>% 
  group_by(Subject.Number, organism) %>% 
  mutate(subject_exposure_perstrain = mean(est_relative_abundance_panel_alt)) %>%
  ungroup() %>%
  distinct(Subject.Number, organism, .keep_all = TRUE) %>%  
  select(Subject.Number, organism, TRT.norm,  rec.diagnosis.Wk8, subject_exposure_perstrain ) %>%
  mutate(subject_exposure_binary = if_else(subject_exposure_perstrain>0, 1, 0))
  
  
 
exposures <- exposures_relative
exposures$TRT.norm = factor(exposures$TRT.norm,
                                         levels = c("Placebo", "VE303 Low Dose"  , "VE303 High Dose"))
textsize <- 14

##############


extfig2D <- ggplot(exposures, 
       aes(x=organism, y=((subject_exposure_perstrain*100) + 0.01), fill=TRT.norm, color=TRT.norm)) +
  geom_boxplot(alpha=.2,  outlier.shape = NA,width = 0.7,show.legend = TRUE ) +
  geom_point(pch = 21, size = 2.5, alpha = 0.5, position = position_jitterdodge() , show.legend = FALSE ) +
  theme_bw() +
#  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
#  facet_grid(.~TRT.norm, scales = "free_x", space = "free") +
  scale_fill_manual(values = coh.cols) +
  scale_color_manual(values = coh.cols) +
  theme(axis.text.x = element_text(size = textsize, angle = 45, hjust = 1, vjust =1), 
        axis.title.x = element_text(size = textsize ),
        plot.title = element_text(size = textsize ),
        axis.text.y = element_text(size = textsize, hjust = 1, vjust =0.5) ,     
        axis.title.y = element_text(size = textsize, hjust = 0.5, vjust = 1) ,
        strip.text = element_text(size = textsize)) +
 # scale_x_discrete(labels = c("Placebo" = "Placebo" , "VE303 Low Dose" = "VE303\nLow Dose",
#                              "VE303 High Dose" = "VE303\nHigh Dose" )) +
  scale_y_log10( ) +
  labs(title= "", y="Mean exposure per subject, %", x="", 
       color="Treatment", fill="Treatment" )
ggsave(filename = here(results, paste(Sys.Date(), taxplot,"VE303 exposure per subject_EXT Figure2D.pdf", sep = " ")), 
       width = 13, height = 5, dpi = 800 )
##############

 
extfig2D

```
# Plots total VE303 abundance and individual strain abundance over time: Figure 1B, 1C 
# Plots diversity over time: Figure 5A
```{r}
ve303.dnameta <- ve303.dnameta_plot.alt

source("plot.clean.abund.ve303_det_response.R")
#source("plot.abund.ve303_det_response.R")



```

 

 
#Plot for VE303 Research MS
#Combines panel to create Figure 1
```{r}
times <- "2wk"
textsize = 16

abuntime1 <- abuntime + theme(axis.title = element_blank(), legend.position = "none") + labs(title= NULL)
dettime1 <- dettime + theme(legend.position = "none") +
    theme(axis.text.x = element_blank(), axis.title.x =  element_blank(), axis.title = element_blank()) + 
  labs(title= NULL)

abuntime_strains1 <- abuntime_strains_long + theme(axis.title = element_blank(), legend.position = "none") + labs(title= NULL)

patch1 <- wrap_plots(dettime1/abuntime1 ) + theme(plot.tag = element_text(size = 18))
 # plot_annotation(tag_levels = 'A')
ggsave(filename = here(results, paste(Sys.Date(),taxplot, "Research MS PK Figure 2AB.png", 
                                      sep = " ")), width = 8, height = 10, dpi = 800)



labels1 <- paste0("VE303-08 Colonization = 0")
labels2 <- paste0("VE303-08 Colonization > 0")

panel3_clin_median_ve303_08_eventfree1 <- panel3_clin_median_ve303_08_eventfree + labs(title= "Early Colonization: VE303-08")
panel3_clin_median_ve303_08_eventfree2  <- panel3_clin_median_ve303_08_eventfree1 +
  theme(axis.text.x = element_text(size = textsize, angle = 0, hjust = 1, vjust =1), 
        axis.title.x = element_text(size = textsize ),
        plot.title = element_text(size = textsize ),
        axis.text.y = element_text(size = textsize, hjust = 1, vjust =0.5) ,     
        axis.title.y = element_text(size = textsize, hjust = 0.5, vjust = 1) ) +
      scale_colour_manual("",values=c("indianred2","cyan3"),breaks=c("Low Engraft", "High Engraft"), 
                        labels=c(labels1,labels2))+
      scale_fill_discrete(name="Engraftment",labels=c(labels1, labels2)) 
 

panel_1D  <- panel1D + 
  theme(axis.title = element_blank(), axis.title.x = element_text(size = 14),
                             legend.title = element_blank(), legend.position = c(0.83, 0.9),
                             legend.background=element_rect(fill = alpha("white", 0.0))) +
       ylim(0,100) 


patchd <- patch1 + abuntime_strains1 + panel_1D + panel3_clin_median_ve303_08_eventfree2 +
  plot_layout(heights =  c(2, 1))


patchd + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 18))

ggsave(filename = here(results, paste(Sys.Date(),taxplot, times,"Research MS PK Figure 1.png", 
                                      sep = " ")), width = 16, height = 16, dpi = 800)
```

 


## Various LME Models for Shannon Diversity: Model with Screening + Abx type as covariate

```{r}


timeplot <- "With Screening"
abxplot <- "With Abx Cov"

vvshann_lme <- vvtest5 %>% select(Sample.ID, Subject.Number, Visit.Name ,Day.in.treatment ,TRT.norm, rec.diagnosis.Wk8, shannon_diversity, species_richness, Abx.group) %>%
  arrange(TRT.norm, rec.diagnosis.Wk8,Subject.Number, Day.in.treatment ) %>%
  filter(Day.in.treatment < 30 & (!is.na(Visit.Name)) )

vvshann_lme$Abx.group = factor(vvshann_lme$Abx.group, levels = c("Vancomycin", "Fidaxomicin"  , 
                                                                       "Metronidazole", "Other"))

vvshann_lme$TRT.norm = factor(vvshann_lme$TRT.norm, levels = c("Placebo", "VE303 Low Dose"  , "VE303 High Dose"))

vvshann_lme$Visit.Name = factor(vvshann_lme$Visit.Name, levels = c("Screening","Day 1"  , "Day 7" , 
                                                               "Day 14" ,  "Day 28", "Day 56" , 
                                                               "Day 168"))

vvshann_lme$Visit.Name = factor(vvshann_lme$Visit.Name, levels = c( "Screening", "Day 1"  , "Day 7" , 
                                                               "Day 14" ,  "Day 28", "Day 56" , 
                                                               "Day 168"))
 
shannon_model_rec <- lmer(shannon_diversity ~ rec.diagnosis.Wk8 + Abx.group +  (1 |  Subject.Number), vvshann_lme, REML = F)


shannon_model_full <- lmer(shannon_diversity ~ TRT.norm*rec.diagnosis.Wk8*Visit.Name + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)
 
shannon_model_random <- lmer(shannon_diversity ~ (1 | Subject.Number), vvshann_lme, REML = F)
(AIC(shannon_model_random) - AIC(shannon_model_full) )*log2(exp(1))
 

shannon_model_full_trt_visit <- lmer(shannon_diversity ~ TRT.norm*Visit.Name + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

shannon_model_full_trt_rec <- lmer(shannon_diversity ~ TRT.norm*rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)
 

print("Model for Response and Treatment Interaction: Sets Bad Outcome as reference interaction level. Shows coefficient for good vs bad in each treatment")
shannon_model_trt_rec1 <- lmer(shannon_diversity ~ TRT.norm + TRT.norm:rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

print("Model for Response and Treatment Interaction: Sets Placebo as reference interaction level. Shows coefficient for treatment vs placebo given a response group")
shannon_model_trt_rec2 <- lmer(shannon_diversity ~ rec.diagnosis.Wk8 + TRT.norm:rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)


print("Model for Visit and Treatment Interaction: Sets Screening as reference interaction level. Shows coefficient for diversity recovery in each treatment")
shannon_model_trt_visit1 <- lmer(shannon_diversity ~ TRT.norm + TRT.norm:Visit.Name  + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

print("Model for Visit and Treatment Interaction: Sets Placebo as reference interaction level. Shows coefficient for treated versus placebo at each timepoint")
shannon_model_trt_visit2 <- lmer(shannon_diversity ~ Visit.Name + TRT.norm:Visit.Name  + Abx.group + (1 | Subject.Number), vvshann_lme, REML = F)

####################################################################################

save_result <- list(change_t_pertrt = summary(shannon_model_trt_visit1)$coefficients,
                    change_trt_overtime = summary(shannon_model_trt_visit2)$coefficients)

save_results <- lapply(save_result, function(x){
  x <- as.data.frame(x)
  x <- cbind(rownames(x) , x )
  x <- rename(x, contrasts = `rownames(x)`)
  x})


## Writes LME coefficients and pvalues to multi sheet excel file ###
writexl::write_xlsx(save_results, here(results, paste(Sys.Date(),  "LME Shannon_Table S5.xlsx", sep=" ")) )
 

```

 
## Various LME Models for Gini-Simpson: Model with Screening + Abx type as covariate

```{r}
 

timeplot <- "With Screening"
abxplot <- "With Abx Cov"

vvshann_lme <- vvtest5 %>% select(Sample.ID, Subject.Number, Visit.Name ,Day.in.treatment ,TRT.norm, rec.diagnosis.Wk8, shannon_diversity, gini_simpson_index,species_richness, Abx.group, Screening_abx_check) %>%
  arrange(TRT.norm, rec.diagnosis.Wk8,Subject.Number, Day.in.treatment ) %>%
  filter(Day.in.treatment < 70 & (!is.na(Visit.Name)) ) 
#  filter(Screening_abx_check == "True Screening")

vvshann_lme$Abx.group = factor(vvshann_lme$Abx.group, levels = c("Vancomycin", "Fidaxomicin"  , 
                                                                       "Metronidazole", "Other"))

vvshann_lme$TRT.norm = factor(vvshann_lme$TRT.norm, levels = c("Placebo", "VE303 Low Dose"  , "VE303 High Dose"))

vvshann_lme$Visit.Name = factor(vvshann_lme$Visit.Name, levels = c("Screening","Day 1"  , "Day 7" , 
                                                               "Day 14" ,  "Day 28", "Day 56" , 
                                                               "Day 168"))

vvshann_lme$Visit.Name = factor(vvshann_lme$Visit.Name, levels = c( "Screening", "Day 1"  , "Day 7" , 
                                                               "Day 14" ,  "Day 28", "Day 56" , 
                                                               "Day 168"))


 
shannon_model_rec <- lmer(gini_simpson_index ~ rec.diagnosis.Wk8 + Abx.group +  (1 |  Subject.Number), vvshann_lme, REML = F)


shannon_model_full <- lmer(gini_simpson_index ~ TRT.norm*rec.diagnosis.Wk8*Visit.Name + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)
#summary(shannon_model_full)
#drop1(shannon_model_full,test="Chisq")

shannon_model_random <- lmer(gini_simpson_index ~ (1 | Subject.Number), vvshann_lme, REML = F)
(AIC(shannon_model_random) - AIC(shannon_model_full) )*log2(exp(1))
 

shannon_model_full_trt_visit <- lmer(gini_simpson_index ~ TRT.norm*Visit.Name + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

shannon_model_full_trt_rec <- lmer(gini_simpson_index ~ TRT.norm*rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)
 

print("Model for Response and Treatment Interaction: Sets Bad Outcome as reference interaction level. Shows coefficient for good vs bad in each treatment")
shannon_model_trt_rec1 <- lmer(gini_simpson_index ~ TRT.norm + TRT.norm:rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

print("Model for Response and Treatment Interaction: Sets Placebo as reference interaction level. Shows coefficient for treatment vs placebo given a response group")
shannon_model_trt_rec2 <- lmer(gini_simpson_index ~ rec.diagnosis.Wk8 + TRT.norm:rec.diagnosis.Wk8 + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)


print("Model for Visit and Treatment Interaction: Sets Screening as reference interaction level. Shows coefficient for diversity recovery in each treatment")
shannon_model_trt_visit1 <- lmer(gini_simpson_index ~ TRT.norm + TRT.norm:Visit.Name  + Abx.group + (1 |  Subject.Number), vvshann_lme, REML = F)

print("Model for Visit and Treatment Interaction: Sets Placebo as reference interaction level. Shows coefficient for treated versus placebo at each timepoint")
shannon_model_trt_visit2 <- lmer(gini_simpson_index ~ Visit.Name + TRT.norm:Visit.Name  + Abx.group + (1 | Subject.Number), vvshann_lme, REML = F)

####################################################################################

save_result <- list( 
                    change_t_pertrt  = summary(shannon_model_trt_visit1)$coefficients,
                    change_trt_overtime  = summary(shannon_model_trt_visit2)$coefficients )

save_results <- lapply(save_result, function(x){
  x <- as.data.frame(x)
  x <- cbind(rownames(x) , x )
  x <- rename(x, contrasts = `rownames(x)`)
  x})

save_results

## Writes LME coefficients and pvalues to multi sheet excel file ###
writexl::write_xlsx(save_results, here(results, paste(Sys.Date(),   "LME Simpson Table S5.xlsx", sep=" ")) )

```

 

 
